Bayesian Clustering and Topic Discovery: \\ Adventures with Gene Expression Data

### 2017-05-11 finalreport/paper.tex
start mm

% For figures
\usepackage{graphicx} % more modern
\usepackage{subfigure} 
\usepackage{amsfonts}

@@

\section{Dirichlet Mixture Model} 
\label{mmsec}
\subsection{Model Description} 
\subsection{Implementation} 
\subsection{Experiments}  
- comparison with pe'er paper

@@

@@

\subsection{Breakdown of Work}
The implementation(s) and experiments involving LDA, Dirichlet Mixtures, and Dynamic Time Model was completed by SK (Section~\ref{ldasec}, Section~\ref{mmsec}, and Section~\ref{dtmsec}). KY completed the implementations and experiments involved Integrated Topic-Clustering, Non-Parametric models, and the shared gene enrichment test implementation (Section~\ref{intsec} and Section~\ref{nonparametricsec}). The authors contributed equally in this work.

\subsection{Latent Dirichlet Allocation}
\label{ldaappendix}
For our Gibb's sampler, we had a fixed burn-in of number of samples (200), a fixed number of sampling iterations after that (500). We didn't extensively explore varying these values, but trying out significantly more iterations (700) didn't seem to change the topic's word distributions significantly. There was one sampling chain on each of four cores.

@@

\subsection{Dynamic Topic Model}
\label{dtmappendix}
Here is a description of the generative process for our dynamic topic model:
\begin{enumerate}
    \item Draw a new topic composition hyperparameter \\ $\beta \leftarrow \mathbb{N}(\beta_{t-1}, \sigma I^2)$
    \item Draw a new document composition hyperparameter \\ $\alpha \leftarrow \mathbb{N}(\alpha_{t-1}, \delta I^2)$
    \item For each document: 
    \begin{enumerate}
        \item Draw a new document topic distribution \\ $\theta \leftarrow \pi(\mathbb{N}(\alpha_t))$
        \item For every word in the document:
        \begin{enumerate}
            \item Draw the word topic assignment \\ $Z \leftarrow$ Mult$(\theta)$
            \item Draw the word identity \\ $Z \leftarrow$ Mult$(\pi(\beta_{t,z}))$
        \end{enumerate}
    \end{enumerate}
\end{enumerate}
Here, $\pi(x_i)$ is the softmax function $\frac{exp(x_i)}{\sum_k exp(x_k)}$ \cite{dtm}.

@@

>>>>

% For figures
\usepackage{graphicx} % more modern
\usepackage{booktabs}
\usepackage{subfigure} 
\usepackage{amsfonts}

@@

\section{Mixture Model} 
\label{mmsec}
\subsection{Model Description} 
The second question we wanted to address
- comparison with pe'er paper
\subsection{Implementation} 
\subsection{Experiments}  
- 500 sampling iterations per run
\begin{table}[h]
    \centering
    \label{mmtable}
    \begin{tabular}{@{}ll@{}}
        \toprule
        \# of Clusters & \# of Clusters Used \\ \midrule
        10                 & 3                                                         \\
        25                 & 9                                                         \\
        50                 & 16                                                        \\
        75                 & 18                                                        \\
        100                & 20                                                        \\ \bottomrule
    \end{tabular}
    \caption{The number of clusters defined in the model appears to be more than the actual number of unique clusters assigned to the data points after posterior estimation. The estimated true number of categories is slightly above 10.}
\end{table}

@@

\section{Breakdown of Work}
The implementation(s) and experiments involving LDA, Gaussian Mixtures, and Dynamic Time Model was completed by SK (Section~\ref{ldasec}, Section~\ref{mmsec}, and Section~\ref{dtmsec}). KY completed the implementations and experiments involved Integrated Topic-Clustering, non-parametric models, (Section~\ref{intsec} and Section~\ref{nonparametricsec}) and the shared gene enrichment test implementation. The authors contributed equally in this work.

@@

\subsection{Latent Dirichlet Allocation}
\label{ldaappendix}

This section includes Figure~\ref{fig:pathways} on the next page.

For our Gibb's sampler, we had a fixed burn-in of number of samples (200), a fixed number of sampling iterations after that (500). We didn't extensively explore varying these values, but trying out significantly more iterations (700) didn't seem to change the topic's word distributions significantly. There was one sampling chain on each of four cores.

@@

\subsection{Dynamic Topic Model}
\label{dtmappendix}

This section includes Figure~\ref{fig:pathwaysdtm} on the next page.

Here is a description of the generative process for our dynamic topic model:
\begin{enumerate}
    \item Draw a new topic composition hyperparameter \\ $\beta \leftarrow \mathbb{N}(\beta_{t-1}, \sigma I^2)$
    \item Draw a new document composition hyperparameter \\ $\alpha \leftarrow \mathbb{N}(\alpha_{t-1}, \delta I^2)$
    \item For each document: 
    \begin{enumerate}
        \item Draw a new document topic distribution \\ $\theta \leftarrow \pi(\mathbb{N}(\alpha_t))$
        \item For every word in the document:
        \begin{enumerate}
            \item Draw the word topic assignment \\ $Z \leftarrow$ Mult$(\theta)$
            \item Draw the word identity \\ $Z \leftarrow$ Mult$(\pi(\beta_{t,z}))$
        \end{enumerate}
    \end{enumerate}
\end{enumerate}
Here, $\pi(x_i)$ is the softmax function $\frac{exp(x_i)}{\sum_k exp(x_k)}$ \cite{dtm}.

@@

\subsection{Gaussian Mixture Model}

For our Gibb's sampler, we had a fixed burn-in of number of samples (200), a fixed number of sampling iterations after that (500). For time reasons, we didn't extensively explore varying these values. Figure~\ref{fig:llmm} describes the log-likelihood evaluation of the estimated posteriors.

\begin{figure*}
    \centering
    \includegraphics[width=1\textwidth]{figs/llmm}
    \caption{Log-likelihood of held-out test set, on three seperate runs of Gibbs sampling procedure, described above}
    \label{fig:llmm}
\end{figure*}

\label{mmappendix}
